---
title: "Hakai Institute Juvenile Salmon Data Package Quality Report"
author: "Brett Johnson"
date: '`r date()`'
output:
 html_document:
   theme: cosmo
   code_folding: hide
   toc: true
   toc_float: true
   number_sections: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = FALSE, error = TRUE)
library(tidyverse)
library(lubridate)
library(here)

survey_data <- read_csv(here("data", "survey_data.csv"))
seine_data <- read_csv(here("data", "seine_data.csv"))
survey_seines <- right_join(survey_data, seine_data, by = "survey_id")

fish_data <- right_join(read_csv(here("data","fish_lab_data.csv"), guess_max = 20000), read_csv(here("data", "fish_field_data.csv"), guess_max = 20000), by = "ufn")
survey_seines_fish <- right_join(survey_seines, fish_data, by = "seine_id") 

```


```{r}
n_fish_collected <- nrow(survey_seines_fish)
n_fish_dissected <- nrow(survey_seines_fish %>% drop_na(date_processed))
n_na_seine_id <- n_fish_collected - nrow(survey_seines_fish %>% drop_na(seine_id))

# count number of NAs for every column and provide as a percent
n_nas <- sapply(survey_seines_fish, function(x) sum(is.na(x)) / n_fish_collected * 100)
# It might be useful to consider which columns should have zero NAs and select those out to have fewer variables to read. Which ones should be 100% filled in? ufn, semsp_id, survey_id, seine_id....

#TODO: Plot or table of species catch broken out by species (pie or stacked barplot)
#TODO: Calculate condition of fish to find outliers in length and weight
```



# Data Quality

## survey_seines joined

```{r}

std_survey_seines_w_catch <- survey_seines %>%
   filter(fish_retained == "yes",
          survey_type == "standard",
          collection_protocol == "SEMSP")

no_lat <- std_survey_seines_w_catch %>% filter(is.na(lat)) %>% 
   select(survey_id, seine_id, survey_type, collection_protocol, fish_retained, lat, long)

# Uncomment this to find out which seines have no associated location data
# knitr::kable(no_lat, caption = "Table. Seines with no associated location information")
```


```{r}

no_fish_loc <- survey_seines_fish %>% filter(survey_type == "standard",
                                     is.na(survey_seines_fish$lat))

```

We've captured `r n_fish_collected` fish and dissected `r n_fish_dissected`

`r nrow(no_fish_loc)` fish have no location information associated with them.

## Sample results

### Sockeye Diets

```{r}
sockeye_diets <- read_csv(here("data", "sample_results", "sockeye_diets.csv"))

test_join <- anti_join(sockeye_diets, fish_data)
```

We have `r length(unique(sockeye_diets$ufn))` diet results for juvenile sockeye

# Relational Joins

Do all seines have a survey to match to?

```{r}
#antijoin should yield zero
seines_match_surveys <-nrow(anti_join(seine_data, survey_data))

ifelse(seines_match_surveys == 0, "yes, seines match surveys", "error: no, some seines don't have matching surveys")
```

Do all lab fish have a matching field fish?

```{r}
fish_field_data <- read_csv(here("data", "fish_field_data.csv"), guess_max = 20000)
fish_lab_data <- read_csv(here("data", "fish_lab_data.csv"), guess_max = 20000)

lab_match_field <- nrow(anti_join(fish_lab_data, fish_field_data))

ifelse(lab_match_field == 0, "yes, all lab fish have a matching field fish", "error: no, some lab fish do not have a mathing field fish")
```

Do all fish have a matching seine?

```{r}

n_umatched_fish <- nrow(anti_join(fish_data, seine_data, by = 'seine_id'))

ifelse(lab_match_field == 0, "yes, all fish have a matching seine they came from", "error: no, not all fish have a seine associated with it")
```

Do all sealice field samples match to a fish?

```{r}
sealice_field <- read_csv(here("data", "sealice_field.csv"))

n_missed_field_lice <- nrow(anti_join(sealice_field, fish_data))

ifelse(n_missed_field_lice == 0, "yes, all sealice counts belong to a fish we collected", "error: some sealice counts don't match to a fish in our database")
```

Do all finescale sealiced fish have a match in our fish data table?

```{r}
sealice_lab_fs <- read_csv(here("data", "sample_results", "sealice_lab_fs.csv"))

fine_loused_matched_fish <- nrow(anti_join(sealice_lab_fs, fish_data, by = "ufn"))

ifelse(fine_loused_matched_fish == 0, "yes, all finescale sealiced fish have a match in our fish data table", "error: no")
```

Do all lab mot sealice results have a match to a fish in our fish data table?

```{r}
sealice_lab_mot <- read_csv(here("data", "sealice_lab_mot.csv"))

n_unmatched_lab_mot <- nrow(anti_join(sealice_lab_mot, fish_data, by = "ufn"))

ifelse(n_unmatched_lab_mot == 0, "yes, all lab mot selice results have a match to a fish", "error: no")
```

# Locations

Visually check to see if any lat and longs are on land and then hover over the site to identify the problematic seine_id.

```{r}
library(leaflet)
library(ggridges)
library(leaflet.esri)
map_data <- survey_seines %>% select(lat, lon = long, site = seine_id) %>% 
   mutate(type = "seine") %>% 
   group_by(site) %>% 
   drop_na(lat) %>% 
      ungroup()
 
pal <- colorFactor(c("navy", "red"), domain = c("seine", "ocgy"))
    
    leaflet() %>% 
      setView(lng = -125.7, lat = 50.4, zoom = 7) %>% 
      addCircleMarkers(data = map_data, color = ~pal(type),
                       stroke = FALSE, fillOpacity = 0.2,
                    label = ~as.character(site),
                       labelOptions = labelOptions(noHide = F, direction = "bottom")
      ) %>%
      addProviderTiles(providers$Esri.OceanBasemap,  options = providerTileOptions(opacity = 1)) %>%
      addEsriTiledMapLayer(url = "https://ags.hakai.org:6443/arcgis/rest/services/AGOL_basemaps/Marine_charts/MapServer") %>%  
      addProviderTiles(providers$CartoDB.VoyagerOnlyLabels) %>% 
      addMiniMap()
```

