---
title: "Hakai Institute Juvenile Salmon Data Package Quality Report"
author: "Brett Johnson"
date: '`r date()`'
output:
 html_document:
   theme: cosmo
   code_folding: hide
   toc: true
   toc_float: true
   number_sections: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = FALSE, error = TRUE)
library(tidyverse)
library(lubridate)
library(here)

sites <- read_csv(here("data","sites.csv"))
survey_data <- read_csv(here("data", "survey_data.csv"))%>% 
  left_join(select(sites,
                   site_id,
                   site_name,
                   region,
                   zone),
            by = "site_id")
seine_data <- read_csv(here("data", "seine_data.csv"))
survey_seines <- right_join(survey_data, seine_data, by = "survey_id")

fish_data <- right_join(read_csv(here("data","fish_lab_data.csv"), guess_max = 20000), read_csv(here("data", "fish_field_data.csv"), guess_max = 20000), by = "ufn")
survey_seines_fish <- right_join(survey_seines, fish_data, by = "seine_id") 

```

# Sea lice

Sea lice is probably the most complicated set of data to combine largely because we used so many different methods. We counted lice in the field under different protocols, and we counted them in the lab under different protocols as well. Some methods produce a very fine level of resolution when it comes to species and stage ID whereas other methods are more coarse and only focus on larger 'motile' sealice that can be seen easier. 

Our goal is to create a time series of motile (adult, not attached and easy to see and ID) sea lice counts that use methods that are reasonably similar enough to warrant combining observations to make interannual comparisons. The lowest common denominator among all years is that motile sea lice where identified to species. Thus the time series doesn't contain any stage data.

Each year a new table that contains the motile caligus and lep counts should be added to the time series in a similar way to how the other sea lice tables are combined. Alternatively, if the motile lab method stays the same, the data can just be added to JSP Master Data Tables google sheet and then exported as a .csv and saved in this repository.

```{r historic sealice integration}
# This script combines various field and lab sealice count methods, and shouldn't have to be modified between years which is why it's abstracted away into a separate script that  sources in an integrated field_lice table and an integrated lab_lice table. To update the annual time series, just add the current years data to either the lab or field lice table in the code chunk below.
source(here::here("scripts", "time_series_data_wrangle","historic_sealice_integration.R"))
```


```{r current-year sea lice integration}

# In 2020 the stages were not identified in the lab but just counts of motile leps, or caligus  were conducted. So I just downloaded the google sheet manually to jsp-data and import that table here separately and combine with lab_lice before joining with field_lice to produce the time series. In 2021 if the sea lice data are added to the JSP Master Data Tables googlesheet, just change the name of the .csv file when it's exported from googlesheets to indicate 2021 is included in that file as well, and then update the read_csv code below to refelect the name of the file.
sealice_current <- read_csv(here("data", "sample_results", "sealice", "sealice_lab_mots_2020.csv")) %>%
  rename(motile_caligus_lab = cal_count, motile_lep_lab = lep_count) %>% 
  left_join(survey_seines_fish) %>% 
  select(ufn, site_id, species, region, survey_date, motile_caligus_lab, 
         motile_lep_lab)

full_lab_lice <- bind_rows(lab_lice, sealice_current)
rm(lab_lice, sealice_current)


sealice_time_series <- full_join(full_lab_lice, field_lice) %>%
  # with preference to lab ID, combine field and lab ID columns into one column
  mutate(motile_caligus = coalesce(motile_caligus_lab, motile_caligus_field)) %>%
  mutate(motile_lep = coalesce(motile_lep_lab, motile_lep_field)) %>%
  mutate(all_lice = motile_caligus + motile_lep) %>%
  select(ufn,
         survey_date,
         site_id,
         region,
         species,
         motile_lep,
         motile_caligus,
         all_lice) %>%
   gather(`motile_caligus`,
         `motile_lep`,
         `all_lice`,
         key = louse_species,
         value = n_lice) %>%
  mutate(year = year(survey_date)) %>% 
  filter(
    site_id %in% consistent_sites
    ) %>%
  drop_na() %>%
  filter(species %in% c("SO", "PI", "CU"))

rm(full_lab_lice, field_lice)

write_csv(sealice_time_series, here("data", "sealice_time_series.csv"))
saveRDS(sealice_time_series,  here::here("data", "sealice_time_series.RDS"))
```

```{r}
n_fish_collected <- nrow(survey_seines_fish)
n_fish_dissected <- nrow(survey_seines_fish %>% drop_na(date_processed))
n_na_seine_id <- n_fish_collected - nrow(survey_seines_fish %>% drop_na(seine_id))

# count number of NAs for every column and provide as a percent
n_nas <- sapply(survey_seines_fish, function(x) sum(is.na(x)) / n_fish_collected * 100)
# It might be useful to consider which columns should have zero NAs and select those out to have fewer variables to read. Which ones should be 100% filled in? ufn, semsp_id, survey_id, seine_id....

#TODO: Plot or table of species catch broken out by species (pie or stacked barplot)
#TODO: Calculate condition of fish to find outliers in length and weight
```



# Data Quality

## survey_seines joined

```{r}

std_survey_seines_w_catch <- survey_seines %>%
   filter(fish_retained == "yes",
          survey_type == "standard",
          collection_protocol == "SEMSP")

no_lat <- std_survey_seines_w_catch %>% filter(is.na(lat)) %>% 
   select(survey_id, seine_id, survey_type, collection_protocol, fish_retained, lat, long)

# Uncomment this to find out which seines have no associated location data
# knitr::kable(no_lat, caption = "Table. Seines with no associated location information")
```


```{r}

no_fish_loc <- survey_seines_fish %>% filter(survey_type == "standard",
                                     is.na(survey_seines_fish$lat))

```

We've captured `r n_fish_collected` fish and dissected `r n_fish_dissected`

`r nrow(no_fish_loc)` fish have no location information associated with them.

## Sample results

### Sockeye Diets

```{r}
sockeye_diets <- read_csv(here("data", "sample_results", "sockeye_diets.csv"))

test_join <- anti_join(sockeye_diets, fish_data)
```

We have `r length(unique(sockeye_diets$ufn))` diet results for juvenile sockeye

# Relational Joins

Do all seines have a survey to match to?

```{r}
#antijoin should yield zero
seines_match_surveys <-nrow(anti_join(seine_data, survey_data))

ifelse(seines_match_surveys == 0, "yes, seines match surveys", "error: no, some seines don't have matching surveys")
```

Do all lab fish have a matching field fish?

```{r}
fish_field_data <- read_csv(here("data", "fish_field_data.csv"), guess_max = 20000)
fish_lab_data <- read_csv(here("data", "fish_lab_data.csv"), guess_max = 20000)

lab_match_field <- nrow(anti_join(fish_lab_data, fish_field_data))

ifelse(lab_match_field == 0, "yes, all lab fish have a matching field fish", "error: no, some lab fish do not have a mathing field fish")
```

Do all fish have a matching seine?

```{r}

n_umatched_fish <- nrow(anti_join(fish_data, seine_data, by = 'seine_id'))

ifelse(lab_match_field == 0, "yes, all fish have a matching seine they came from", "error: no, not all fish have a seine associated with it")
```

Do all sealice field samples match to a fish?

```{r}
sealice_field <- read_csv(here("data", "sealice_field.csv"))

n_missed_field_lice <- nrow(anti_join(sealice_field, fish_data))

ifelse(n_missed_field_lice == 0, "yes, all sealice counts belong to a fish we collected", "error: some sealice counts don't match to a fish in our database")
```

Do all finescale sealiced fish have a match in our fish data table?

```{r}
sealice_lab_fs <- read_csv(here("data", "sample_results", "sealice_lab_fs.csv"))

fine_loused_matched_fish <- nrow(anti_join(sealice_lab_fs, fish_data, by = "ufn"))

ifelse(fine_loused_matched_fish == 0, "yes, all finescale sealiced fish have a match in our fish data table", "error: no")
```

Do all lab mot sealice results have a match to a fish in our fish data table?

```{r}
sealice_lab_mot <- read_csv(here("data", "sealice_lab_mot.csv"))

n_unmatched_lab_mot <- nrow(anti_join(sealice_lab_mot, fish_data, by = "ufn"))

ifelse(n_unmatched_lab_mot == 0, "yes, all lab mot selice results have a match to a fish", "error: no")
```

# Locations

Visually check to see if any lat and longs are on land and then hover over the site to identify the problematic seine_id.

```{r}
library(leaflet)
library(ggridges)
library(leaflet.esri)
map_data <- survey_seines %>% select(lat, lon = long, site = seine_id) %>% 
   mutate(type = "seine") %>% 
   group_by(site) %>% 
   drop_na(lat) %>% 
      ungroup()
 
pal <- colorFactor(c("navy", "red"), domain = c("seine", "ocgy"))
    
    leaflet() %>% 
      setView(lng = -125.7, lat = 50.4, zoom = 7) %>% 
      addCircleMarkers(data = map_data, color = ~pal(type),
                       stroke = FALSE, fillOpacity = 0.2,
                    label = ~as.character(site),
                       labelOptions = labelOptions(noHide = F, direction = "bottom")
      ) %>%
      addProviderTiles(providers$Esri.OceanBasemap,  options = providerTileOptions(opacity = 1)) %>%
      addEsriTiledMapLayer(url = "https://ags.hakai.org:6443/arcgis/rest/services/AGOL_basemaps/Marine_charts/MapServer") %>%  
      addProviderTiles(providers$CartoDB.VoyagerOnlyLabels) %>% 
      addMiniMap()
```

